### 2014-04-15 doc/uist14/interaction.tex
bjoern's pass

\section{Interaction Models}

In this section, we first describe how head orientation-based interaction works from the users' point of view (Figure~\ref{fig:intearaction}).

{\bf Scan:} The user first scans the environment to locate the position of the target. During this stage, Glass will constantly sends out IR signal, and the target offers immediate feedback when it receives the IR.

{\bf Coarse Selection:} The user confirms his desire to connect to a target by tapping on the Glass touchpad. At this time, Glass collects the replies from targets about the IR signal reception. If there is only a single target in IR range, then it's considered as a successful selection. However, in a relative dense environment where multiple targets are within range, the user needs to perform a fine selection.

{\bf Fine selection:} When disambiguation is needed, the user needs to make an explicit selection among the targets that within his view range. There are multiple ways to achieve it -- either from Glass touchpad that controls a UI list or using the head motion sensor on Glass. We will leave a thorough and complete discussion later. Once the user confirms the current selection, he could tap to complete the acquisition. Since this part is effectively disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/interactionModel.pdf}
\caption{Four different states during the head-orientation targeting.}
\label{fig:interaction}
\end{figure}

From the flow of interaction, we learned that the overall target acquisition time can be broken down into the following pieces:
\begin{equation}
t_{total}=t_{scan}+t_{coarse}+t_{disambiguation}
\end{equation}

Our goal of this paper is to minimize the overall target selection time. In the following few sections, we will describe our iterative design process during which we have explored the design space.

>>>>

%!TEX root = uist14.tex
\section{IR for head orientation sensing}
\subsection{Hardware}
A central hypothesis of this paper is that the area cursor paradigm~\cite{kabbash1995prince} is well matched to head orientation selection. Head orientation input is imprecise because it does not capture eye movement and has to rely on a low-bandwidth muscle group. Therefore, point selection techniques like laser pointers are not appropriate.

Infrared emitters are a good technology match since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible. LEDs with many different fields of view are commercially available. \bjoern{say something about receivers and how point to point connections enable us to leverage head orientation without needing a map.}

Another benefit is the low cost and easy integration of both emitters and detectors - detectors often have support for common IR protocol decoding built in and cost cents, making it economically feasible to add detectors to dozens or hundreds of devices in an environment.

Our augmentation to Google Glass adds a \bjoern{add modelnumber and spec details} LED (see Figure~\ref{fig:glass}. In the following description, we refer to Glass because the term is succinct; but  other suitable head-worn computing devices could be substituted.

\bjoern{need to mention that all the target nodes also have to be able to talk to each other and to glass outside the IR channel. We use XBees for this purpose.}

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/tube_from_top_annotated.pdf}
\caption{Google Glass augmented with a repositionable IR holder.}
\label{fig:glass}
\end{figure}

\subsection{Interaction Model}
From the user's perspective, interaction with \systemname proceeds in two stages (Figure~\ref{fig:interaction}):

{\bf Scan:} The user first scans the environment to locate the position of the target. During this stage, Glass constantly sends out IR signals, and  targets offer immediate feedback when they receive a signal. The user confirms his desire to connect to a target by tapping on the Glass touchpad. Glass collects the responses from targets that have received IR reception (we use an XBee-based backchannel for this traffic). If there is only one single target in IR range, it is automatically selected. However, in a dense environment where multiple targets are within range, the user needs to perform a refinement.

{\bf Refine:} When disambiguation is needed, the user needs to make an explicit selection among the targets that within his view range. We have designed multiple refinement mechanisms -- all enable the user to select one from a subset of targets. The user confirms a current selection a tap. Since the purpose of this stage is to disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage. Again, a tap confirms a selection.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/interactionModel2.pdf}
\caption{Four different states during the head-orientation targeting.}
\label{fig:interaction}
\end{figure}

The overall target acquisition time thus depends on scan and refine times; the probability that refinement is needed; and the time to commit an action (tap):
\begin{equation}
t_{total}=t_{scan}+P(refine)*t_{refine}+t_{commit}
\end{equation}

In the following sections, we describe our iterative design and evaluation process to minimize the overall target selection time.

### 2014-04-15 doc/uist14/iteration1.tex
bjoern's pass

\section{Iteration 1: IR vs. List}
\subsection{Problem Statement and Goal}
Our system design uses IR as the beam cursor for physical target acquisition because the area coarse selection filters out the targets outside of the user's view range. This help during the fine selection time because there is a smaller set of targets to select from. The opposite is to have a complete list of all targets when there is no coarse selection and the disambiguation is always needed. This brought up the first question that we asked ourselves -- {\em how much does using IR help in reducing the selection time}. To answer that question and quantify the benefit of having IR for the coarse selection, we carried out our first target acquisition study which compares the {\em IR selection} and {\em list selection}.

\subsection{Technique}

We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}). Each of them has a number as ID and a letter representing the name. There are 14 participants from our institution for this study. 13 had never used Glass before, therefore we offered a Google Glass crash course before the experiment for them to get familiar with Glass gestures. 4 of them wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.

@@

In the within-subject study, half of them performed {\em IR selection} first and the other half did {\em list selection} first. For each selection condition, we conducted 15 target selection by randomly choosing from all the targets. During the study, we measure the {\bf target acquisition time} for each target selection, and afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.

@@

\subsection{Evaluation}
Our results indicate that {\em IR selection} is generally better than {\em list selection}. From Figure~\ref{fig:selection-time}A, the average target acquisition time for {\em IR selection} is 6.67 seconds, while for {\em list selection} it took 8.86 seconds. We run Student's t-test on the data and it shows significant difference ($t(279)=-3.81, p=0.00017$).

To further understand if the gain comes from the coarse selection, we splited the data from {\em IR selection} into two parts -- ones that require disambiguation and ones that don't. The result is shown in Figure~\ref{fig:selection-time}B. It shows that it takes 6.40 seconds (on average) to complete the connection if coarse selection has achieved the acquisition, while when multiple devices are in range, the time increases to 9.16 seconds, indicating additional 2.76 seconds for disambiguation. Though it takes significantly longer ($t(19)=-2.7827, p=0.012$ using t-test) in the latter case, they only happen 10\% of total {\em IR selection} trials.

From the above result, we learned that with IR, we can have effectively reduced the overall target acquisition time. And the major gain comes from the fact that {\em coarse selection} helps reduce the chances of entering {\em fine selection} stage.

@@

To further generalize the results, in Figure~\ref{fig:time-vs-list-order}, we show the acquisition time for each individual target in {\em list selection} condition, ordered by their relative position in the list. Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}. From this figure, we can see that once the target's order is larger than 6, the average acquistion time will be larger than the case when disambiguation is needed in {\em IR selection}. Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable. Further more, in real-life scenarios, targets would have longer names than an alphabet letter. This would increase the searching and text comparison time and also reduce the number of targets being shown in the UI screen.

>>>>

\section{Iteration 1: Naive IR}
\subsection{Problem Statement and Goal}
Our initial research question is: {\em Can IR-based targeting reduce the selection time compared to UI list navigation on a head-worn device?}

\subsection{Technique}
In our first implementation, we use IR for scanning as described in the previous section. For the refinement technique, we simply show a list of the subset of targets that have received IR signals on the Glass display and let users swipe to select from that list and tap to confirm.

A natural point of comparison is an interface that does not use any orientation information - it always shows 
a complete list of all targets. We implemented a list view where users swiped forwards and backwards on the Glass touchpad to navigate incrementally through the list. To quantify the benefit of using IR for the scanning step, we carried out a target acquisition study which compares the {\em Naive IR selection} and {\em list selection}.

\subsection{Method}
We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}). Each of them has a number as ID and a letter representing the name. We recruited 14 participants from our institution for this study. 13 had never used Glass before, so we offered a Google Glass tutorial before the experiment for them to get familiar with Glass gestures. Four participants wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle. In this study, the IR LED was fixed and not repositionable as shown in Figure~\ref{fig:glass}.

@@

In the within-subject study, half of the participants performed {\em IR selection} first and the other half used {\em list selection} first. For each selection condition, we conducted 15 target selection by randomly choosing from all the targets. During the study, we measure the {\bf target acquisition time} for each target selection. Afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.

@@

\subsection{Results}
Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds, while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:selection-time}A). Student's t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).

To understand how scanning and refinement contribute to total selection time, we split the data from {\em IR selection} into two parts -- trials that required refinement and ones that did not. It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:selection-time}B). This difference is significant ($t(19)=-2.7827, p=0.012$ using t-test \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}). Because many targets were spaced far apart, refinements were only necessary in 10\% of total {\em IR selection} trials in this study.

@@

To further generalize the results, in Figure~\ref{fig:time-vs-list-order}, we show the acquisition time for each individual target in {\em list selection} condition, ordered by their relative position in the list. Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}. From this figure, we can see that once there are more than 6 targets, the average acquistion time will be larger than IR selection, even if disambiguation is needed. %Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable. Further more, in real-life scenarios, targets would have longer names than an alphabet letter. This would increase the searching and text comparison time and also reduce the number of targets being shown in the UI screen.

In summary, {\em Naive IR} can outperform linear list selection, but the refinement step detracts significantly from the efficiency of the technique.
% learned that with IR, we can have effectively reduced the overall target acquisition time. And the major gain comes from the fact that {\em coarse selection} helps reduce the chances of entering {\em fine selection} stage.

### 2014-04-15 doc/uist14/iteration2.tex
bjoern's pass

\section{Iteration 2: Binary IR vs. Valued IR}
\sean{the term Binary IR and Valued IR can be refined.}
\subsection{Problem Statement and Goal}
With our first study, we learn that using IR to perform {\em coarse selection} can reduce the overall target selection time. A major reason for the improvement is because disambiguation is only needed 10\% of the time. 
%because in many cases there will be only one target in range and fine selection is not needed, and in cases where it’s needed, the list length has been effectively reduced. 
However, if the environment becomes denser, more disambiguation will be needed. And learning from Figure~\ref{fig:selection-time}B, we know the average results when disambiguation is involved is significantly increased. Hence our next research question becomes ``can we improve selection time in a dense environment?''

\subsection{Technique}
Previously we only used IR reception as a signal for identifying potential targets. It's a binary value indicating whether a target has received IR signal or not. To further capture the user's attention, we believe that IR intensity at the receiver side can provide more information. The IR intensity falls when the distance between IR emitter and receiver increases as well as when the receiving angle increases. We have empirically measured the intensity distribution at the receiver side, and it's shown in Figure~\ref{fig:measurement}. \ben{still working on it, not sure how much value we will get out of it. Seems useful to support the argument of the benefit though.}

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/placeholder.pdf}
\caption{placeholder for emperical measurement.}
\label{fig:measurement}
\end{figure}

From the measurement above we learned that the angular difference has a large effect on the intensity readings. In order to utilize the intensity, we have designed and 3D printed a repositional IR holder that can be clipped onto Glass. Thus, different users can adjust the emitter to align with their line of sight. \ben{need to mention in iteration 1 about the crappy tape}.

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/tube_from_top.jpg}
\caption{The repositional IR holder.}
\label{fig:measurement}
\end{figure}

The intensity information is used in two ways:
\begin{enumerate}
\item When multiple targets have received IR signal and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half, which is frequently used because it indicates a 3dB loss in the signal strength.}. Therefore, when there is only one target within line of sight, the IR intensity approach has the same behavior as the previous iteration - no disambiguation is needed. When the environment turns denser, the new design can filter some peripheral targets out, making the chances of entering disambiguation lower.
\item When disambiguation is still needed, meaning that multiple targets have relatively close intensity values, the system sorts the disambiguation list according to the IR intensity, from strongest to weakest. This will help us reduce the $t_{disambiguation}$ significantly if the first one appears matches the actual target.
\end{enumerate}

We seek to quantify the improvement of such design and performed another user study to compare the {\em Binary IR} and {\em Valued IR} approaches. For this study, we aim to discover the performances at a denser environment. Thus, we take the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}). In total, we have 10 users for this study, and each user performs 30 target acquisition tasks for each approach. Similarly to our first within-subject study, half of them perform {\em Binary IR} first and the other half {\em Valued IR} first.

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/study-layout2.jpg}
\caption{The current environment setup \ben{needs annotation and a better photo.}.}
\label{fig:study-layout2}
\end{figure}

@@

\subsection{Evaluation}
Below is a summary of the study result.
\begin{enumerate}
\item Among the 300 trials in {\em Binary IR} condition, there are 225 times when disambiguation is needed. In {\em Valued IR} condition, only 167 out of 300 trials need disambiguation. We perform a Chi-square test on this result and the result is $\chi^2(1) = 24.755$, $p < 0.0001$. This demonstrates that the new approach successfully reduces the chances where disambiguation is needed.
\item In the cases when disambiguation is inevitable, the {\em Valued IR} will sort the list based on the intensity reading. As a result, out of the 167 times when disambiguation is involved, {\em Valued IR} sorts the desired target as the first one in the list 93 times (roughly 55\%). In comparison, for {\em Binary IR}, only 80 out of 219 (35\%) times can the user find the desired target first in the list. Again a Chi-square test demonstrates the statistical difference (with $\chi^2(1) = 15.758$, $p < 0.0001$. 
\end{enumerate}

Such gain can be seen from Figure~\ref{fig:study2} where the overall target acquisition time is obtained. The average time is 4.31 seconds and 3.64 second for {\em Binary IR} and {\em Valued IR}, respectively; and a t-test shows that they are also significantly different $t(555)=3.2945$, $p=0.001$.

@@

However, one side effect that we have observed in this approach is that the {\em Valued IR} sometimes eliminates the desired target during the {\em coarse selection} stage. The rate is not high (13 out of 300 cases, 4.3\%), but this is the trade-off for the overall faster acquisition time.

>>>>

%!TEX root = uist14.tex

\section{Iteration 2: Intensity IR}

\subsection{Problem Statement and Goal}
%With our first study, we learn that using IR to perform {\em coarse selection} can reduce the overall target selection time. A major reason for the improvement is because disambiguation is only needed 10\% of the time. 
%because in many cases there will be only one target in range and fine selection is not needed, and in cases where it’s needed, the list length has been effectively reduced. 
Performance of the {\em Naive IR} technique will degrade as target density in an environment increases, as increased density will require refinement steps. We therefore ask a follow-on research question: ``How might we improve selection time in a dense environment?''

\subsection{Technique}
Previously we only used IR reception as a binary signal for identifying potential targets. We hypothesize that IR intensity at the receiver side can provide more information about the likelihood that a user intended to select a particular target. Received IR intensity falls off with distance between IR emitter and receiver as well as with the angle between emitter and receiver. To measure intensity, we add \bjoern{add model number and characteristics} and replace the IR emitter with \bjoern{modelnumber, characteristics}.
We have empirically measured the intensity distribution at the receiver for this configuration in Figure~\ref{fig:measurement}. \ben{still working on it, not sure how much value we will get out of it. Seems useful to support the argument of the benefit though.}
Our measurements confirm that angular difference has a large effect on the intensity readings.
\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/placeholder.pdf}
\caption{placeholder for empirical measurement.}
\label{fig:measurement}
\end{figure}

The intensity information is used in two ways:
\begin{enumerate}
\item When multiple targets have received IR signals and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half of the ADC resolution, which is frequently used because it indicates a 3dB loss in the signal strength.}. Therefore, when there is only one target within line of sight, the IR intensity approach has the same behavior as the previous iteration - no disambiguation is needed. When the environment turns denser, the new design can filter some peripheral targets out, reducing $P(refine)$, the likelihood of entering the refinement stage.
\item When refinement is still needed, meaning that multiple targets have relatively close intensity values, the system sorts the disambiguation list according to the IR intensity, from strongest to weakest. We hypothesize that this will reduce time $t_{refine}$ significantly by minimizing extra navigation steps. (Ideally, the first list item will match the actual, intended target.)
\end{enumerate}

\subsection{Method}
To quantify the improvement this design we performed a second study to compare the {\em Naive IR} and {\em Intensity IR} approaches. Because we were interested in discovering performance differences in denser environment, we reposition the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}) \bjoern{give measurements in meters}. We recruited 10 participants for this study. Each  user performed 30 target acquisition tasks for each approach. Similarly to our first within-subject study, half of them perform {\em Naive IR} first and the other half {\em Intesity IR} first.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/study-layout2.jpg}
\caption{The current environment setup \ben{needs annotation and a better photo.} \bjoern{yes}.}
\label{fig:study-layout2}
\end{figure}

@@

\subsection{Results}
{\em Intensity IR} reduces the number of trials in which refinement dialogs are needed from 225 of 300 in {\em Naive IR} to 167 of 300 trials. A Chi-square test shows this difference is significant ($\chi^2(1) = 24.755$, $p < 0.0001$). This demonstrates that the new approach successfully reduces the probability that a disambiguation dialog is needed.

In the cases when disambiguation is inevitable, {\em Intensity IR} sorts the list based on the intensity reading, while {\em Naive IR} sorts alphabetically. {\em Intensity IR} reduces the fraction of refinement trails in which additional list navigation is necessary (i.e., the first, already selected element is incorrect).  {\em Intensity IR} sorted the desired target as the first one in the list in 55\% of cases (93 of 167). In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 219\bjoern{why isn't this 225 as above?}). Again, a Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, $p < 0.0001$.

Such gain can be seen from Figure~\ref{fig:study2} where the overall target acquisition time is obtained. \bjoern{avoid passive voice throughout!} The average selection time decreases from 4.31 seconds for {\em Naive IR} to 3.64 second for {\em Intensity IR}. This difference is also significant ($t(555)=3.2945$, $p=0.001$.) \bjoern{for consistency's sake, can we show a box plot as in figure 5?}

@@

One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage. The this error rate (13 out of 300 cases, 4.3\%), is higher than the error rate for {\em Naive IR} \bjoern{insert error rate}. % but this is the trade-off for the overall faster acquisition time.

### 2014-04-15 doc/uist14/iteration3.tex
bjoern's pass

\section{Iteration 3: Head motion for disambiguation}
\label{sec:iteration-3:-head}
\subsection{Problem Statement and Goal}
Though our {\em Valued IR} system already performs well (with 3.71 seconds on average in a relative dense environment), head-orientation information is only used during the {\em coarse selection} stage. In the {\em fine selection} phase, we rely solely on the near-eye display. This has two problems: First, we are projecting a high dimension distribution to a 1-dimension list. Second, this requires the user to switch their focus back and forth between the physical world and the display. This issue motivated us to design an approach that also utilizes orientation information during the {\em fine selection} stage, so that there would be a more direct mapping between the user's gestures and target selection.

@@

\subsection{Technique}
The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of users' head. However, this absolute orientation cannot be directly used in any indoor environment as the user may change his location \achal{Should we use gender neutral language?}. However, as long as the targets are spread around the periphery, their relative orientation doesn't change. This key observation inspires us to investigate a hybrid solution with IR and sensors on Glass.

There are two phases to this technique: learning the relative orientations and intelligently suggesting targets during disambiguation.

@@

In this mode, only one device has a solid indicator light on at a time. The user can then tilt in the direction of the device he wishes to switch to. We calculate the direction of movement using a low-passed history of orientations, and look for devices that lay in that direction. Once a new device is selected, the frame of reference changes, and the user can repeat the process. Note that in reality, this is a completely smooth operation: it is possible to switch between multiple devices in one continuous motion. This is achieved by automatically adjusting the frame of reference to the currently lit device, but also creating a ``sticky" threshold so that the user is not constantly jumping between devices.

@@

\subsection{Evaluation}
\ben{We then need evaluation results here, quantitative and qualitative.}

>>>>

%!TEX root = uist14.tex
\section{Iteration 3: Head motion refinement}
\label{sec:iteration-3:-head}
\subsection{Problem Statement and Goal}
Both{\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step. This has two problems: First, a spatial layout of target is compressed into a uni-dimensional list. List navigation commands thus have poor spatial mapping. Second, our techniques require the user to switch their focus back and forth between the physical world and the display. This issue motivated us to design an approach that also utilizes orientation information during the {\em refinement} stage, to provide a more direct mapping between the user's actions and target selection.

@@

\subsection{Technique}
The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of a user's head. However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location \achal{Should we use gender neutral language?}\bjoern{yes. just did in the previous sentence}. However, as long as the targets are spread around the periphery, their relative orientation doesn't change. This key observation inspires us to investigate a hybrid solution with IR and orientation sensors on Glass.

There are two phases to this technique: learning the relative orientations or targets in a room and intelligently suggesting targets during disambiguation.

@@

In this mode, only one device has a solid indicator light on at a time. \bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} The user can then tilt in the direction of the device he wishes to switch to. We calculate the direction of movement using a low-pass-filtered history of orientations, and look for devices that lie in that direction. Once a new device is selected, the frame of reference changes, and the user can repeat the process. Note that in reality, this is a completely smooth operation: it is possible to switch between multiple devices in one continuous motion. This is achieved by automatically adjusting the frame of reference to the currently lit device, but also creating a ``sticky" threshold so that the user is not constantly jumping between devices. \bjoern{This is called hysteresis or double thresholding.}

\bjoern{how well will this work if the user shifts position within a room?}

@@

\subsection{Evaluation}
\ben{We then need evaluation results here, quantitative and qualitative.}
\bjoern{got to here in my pass}

### 2014-04-15 doc/uist14/relatedwork.tex
bjoern's pass

\subsection{Area Cursors}
A central hypothesis of this paper is that the area selection paradigm is well matched to head orientation input. In GUI area cursors, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. We argue that head orientation pointing has similar challenges (limited pointing performance and accuracy). In all area cursors, the bigger cursor activation area can lead to multiple targets being selected and disambiguation is needed. 
We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets). In the {\em refinement} phase, they adjust so only the intended target will be selected. Many disambiguation techniques are possible for refinement -- this paper describes the tradeoffs between several of them.

>>>>

\subsection{Area Cursors}
In GUI area cursors, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. We argue that head orientation pointing has similar challenges (limited pointing performance and accuracy). In all area cursors, the bigger cursor activation area can lead to multiple targets being selected and disambiguation is needed. 
We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets). In the {\em refinement} phase, they adjust so only the intended target will be selected. Many disambiguation techniques are possible for refinement -- this paper describes the tradeoffs between several of them.
